---
title: "p8105_hw3_2000"
author: Uma Palaniappan, unp2000
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
## Instacart 
```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart") 
```

`r nrow(instacart)` 1384617 rows 
`r ncol(instacart)` 15 columns 

How many aisles are there and which aisles are the most items ordered from?
There are `r nrow(distinct(instacart, aisle))` distinct aisles.

`r nrow(distinct(instacart, product_name))`

```{r}
instacart_new=instacart %>%
    janitor::clean_names() %>%
    group_by(aisle_id, aisle) %>%
    summarize (n = n()) %>%
    arrange(desc(n))
```
The aisles that the most items are ordered from are fresh vegetables (aisle 83, n = 150,609), fresh fruits (aisle 24, n = 150,473), and packaged vegetable and fruits (aisle 123, n = 78,493). 

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.


```{r}
plot_1= filter(instacart_new, n > 10000)
ggplot(
  plot_1, 
  aes(x = reorder(aisle, -n), y = n)) + 
  geom_point(color = 'red') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text = element_text(size = 7.5)) +
  labs(
    title = "Items ordered in each aisle",
    x = "Aisle",
    y = "No. of products"
  )
  
```


Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
library(knitr)
insta_rank_new = instacart %>%
      janitor::clean_names() %>%
      filter(aisle == "baking ingredients"| aisle == "dog food care"| aisle == 
      "packaged vegetables fruits") %>%
      group_by(aisle, product_name) %>%
      summarise(n = n()) %>%
      mutate(product_ranking = min_rank(desc(n))) %>% 
      filter(product_ranking < 4)  %>%
      knitr::kable ()
```
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart_table2=instacart %>%
  filter(product_name == "Pink Lady Apples" | 
    product_name == "Coffee Ice Cream") %>%
    group_by(product_name, order_dow) %>%
    summarize(mean_order_hour= mean(order_hour_of_day)) %>%
    select(product_name, order_dow, mean_order_hour) %>%
    pivot_wider(
      names_from = order_dow,
        values_from = mean_order_hour) %>%
    knitr::kable ()
```

# Problem 2 

First, do some data cleaning:

format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

```{r}
brfss_new=brfss_smart2010 %>%
    janitor::clean_names() %>%
    filter(topic == "Overall Health"|
           response == "Excellent"|
           response == "Very good"|
           response == "Good" |
           response == "Fair"|
           response == "Poor") %>%
   mutate(response = 
            factor(response, labels = c("Poor", "Fair", "Good", "Very good", "Excellent")))%>%
            select(-location_id, -data_value_footnote_symbol, - data_value_footnote)
``` 
In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_new %>%
  filter(year == "2002") %>%
  group_by(locationabbr) %>%
  summarize(
    number_locations = n_distinct(geo_location)) %>%
    filter(number_locations>6) %>%
    arrange(desc(number_locations))
```
In 2002, the states that were observed at 7 or more locations were NC, FL, CT, NJ, MA, PA. 

In 2010, the states that were observed at 7 or more locations were SC, PA, CO, OH, NY, MA, WA, NE, NC, MD, CA, TX, NJ, FL
```{r}
brfss_new %>%
  filter(year == "2010") %>%
  group_by(locationabbr) %>%
  summarize(
    number_locations = n_distinct(geo_location)) %>%
    filter(number_locations>6) %>%
    arrange(desc(number_locations))
```
In 2010, the states that were observed at 7 or more locations were SC, PA, CO, OH, NY, MA, WA, NE, NC, MD, CA, TX, NJ, FL

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
excellent_data = brfss_new %>%
  filter(response == "Excellent") %>%
  group_by(year, locationabbr, response) %>%
  summarise(mean_data = mean(data_value, na.rm = TRUE)) 

brfss_spaghetti =
  excellent_data %>%
  ggplot(aes(x = year, y = mean_data, group = locationabbr)) +
  geom_line(aes(color = locationabbr)) +
  labs(
    title = "Average value over time within 50 states",
    x = "Year",
    y = "Average Data Value"
  )
brfss_spaghetti
```
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
two_panel = brfss_new %>%
  filter(topic == "Overall Health"|
           response == "Excellent"|
           response == "Very good"|
           response == "Good" |
           response == "Fair"|
           response == "Poor",
           year == "2006"|
           year == "2010",
           locationabbr == "NY") 


ggplot(two_panel, aes(x = response, y = data_value, color = response)) + 
  geom_boxplot() +
  facet_grid(. ~year) +
  labs(
    title = "Distribution of Data Values for Responses among locations in NY State",
    x = "Response Type",
    y = "Data Value"
  )

```

# Problem 3 

```{r}
problem3_data = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(day = factor(day, labels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", 
                                      "Saturday", "Sunday")),
         dow = day == "Saturday" | day == "Sunday",
         dow = ifelse(dow == TRUE, "Weekend", "Weekday")) %>%
  select(week,day, dow, everything (),-day_id) 
```
We have 35 obeservations and 1444 variables in the problem 3 dataset which uses five weeks of accelerometer data that was collected on a 63 year-old male with BMI 25. Additionally, we have `r nrow(problem3_data)` rows and `r ncol(problem3_data)` columns. The variables are the week in which the data was collected, the specific day (Monday - Sunday), the day of the week (either Weekday or Weekend), and the activity per minute for 24 hours (activity_1 to activity_1440)


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
problem3_data %>%
  group_by(week, day) %>%
  summarize(sum_activity= sum(activity_1:activity_1440))%>%
  knitr::kable ()
```
The trends that are apparent 

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
ggplot(problem3_data, aes(x = day, fill = sum_activity)) + 
  geom_histogram (bindwidth =2) +
  labs(
    title = "24 hour activity",
    x = "Day of Week",
    y = "Sum Activity"
  )
```

